Modal Logics in the Coq Proof Assistant

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Improve introduction

However, general automated reasoning support for modal logics is still
not as well-developed as for classical logics. Current deduction tools
for modal logics are often limited to propositional, quantifier-free,
fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular
modal logics and their applications \cite{ToDo:StephanMerz}. Recently,
first-order automated deduction techniques based on tableaux, sequent
calculi and connection calculi have been generalized to cope with
modalities directly and a few new provers have been implemented
\cite{JensOtten}.

Another recently explored possibility \cite{J23,B9,J22} is the
embedding of first-order and higher-order modal logics into classical
higher-order logics, for which existing higher-order automated theorem
provers \cite{ToDo:Leo,Satallax} exist.  Embedding is flexible,
because various modal logics can be easily supported by adding their
characteristic axioms.  The embedding approach can also be easily
adapted to support multiple modalities, and it is possible to replace
constant domain quantifiers by varying or cumulative domain
quantifiers.  Moreover, the approach is relatively simple to
implement, because it does not require any modification in the source
code of the higher-order prover. The prover can be used as is, and
only the input files provided to the prover must be especially
encoded. Furthermore, the efficacy and efficiency of the embedding
approach has been confirmed in philosophically interesting and
relevant benchmarks \cite{FormalTheologyRepository}. These qualities
make embedding a convenient approach for \emph{fully automated}
reasoning.

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant. The main goal of this paper is
to study this question. Our answer is positive. The \Coq proof
assistant was chosen because of the authors' greater familiarity with
the tactic language of this system.

@@

In order to illustrate the use of the implemented embedding, we show
here the formalization of Scott's version \cite{Scott} of Gödel's
ontological argument for God's existence (in Section \ref{sec:proof}).
This proof was chosen mainly for two reasons. Firstly, it requires not
only modal operators, but also higher-order quantification. Therefore,
it is beyond reach of specialized propositional and first-order
(modal) theorem provers. Secondly, this argument addresses an ancient
problem in Philosophy and Metaphysics, which has nevertheless received
a lot of attention in the last 15 years, because of the discovery of
the modal collapse \cite{ToDo}. This proof lies in the center of a
vast and largely unexplored application domain for automated and
interactive theorem provers.

>>>>

However, general automated reasoning support for modal logics is still
not as well-developed as for classical logics. Deduction tools
for modal logics are often limited to propositional, quantifier-free,
fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular
modal logics and their applications \cite{ToDo:StephanMerz}; 
first-order automated deduction techniques based on tableaux, sequent
calculi and connection calculi have only recently been generalized 
and implemented in a few new provers able to directly cope with
modalities \cite{JensOtten}.

Another recently explored possibility \cite{J23,B9,J22} is the
embedding of first-order and even higher-order modal logics into classical
higher-order logics, for which existing higher-order automated theorem
provers \cite{ToDo:Leo,Satallax} exist.  Embedding is flexible,
because various modal logics can be easily supported by adding their
characteristic axioms.  The embedding approach can also be easily
adapted to support multiple modalities, and it is possible to replace
constant domain quantifiers by varying or cumulative domain
quantifiers.  Moreover, the approach is relatively simple to
implement, because it does not require any modification in the source
code of the higher-order prover. The prover can be used as is, and
only the input files provided to the prover must be especially
encoded. Furthermore, the efficacy and efficiency of the embedding
approach has been confirmed in philosophically interesting and
relevant benchmarks \cite{FormalTheologyRepository}. These qualities
make embedding a convenient approach for \emph{fully automated}
reasoning.

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant such as \Coq\footnote{The \Coq
proof assistant was chosen merely because of the authors' greater
familiarity with the tactic language of this system. Nevertheless, the
techniques presented and discussed here are likely to be useful for
other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
\texttt{HOL-Light} \cite{HOLLight}, as well.}. The main goal of this
paper is to study this question. Our answer is positive.

@@

In order to illustrate the use of the implemented embedding, we show
here the formalization of Scott's version \cite{Scott} of Gödel's
ontological argument for God's existence (in Section \ref{sec:proof}).
This proof was chosen mainly for two reasons. Firstly, it requires not
only modal operators, but also higher-order quantification. Therefore,
it is beyond the reach of specialized propositional and first-order
(modal) theorem provers. Secondly, this argument addresses an ancient
problem in Philosophy and Metaphysics, which has nevertheless received
a lot of attention in the last 15 years, because of the discovery of
the modal collapse \cite{ToDo}. This proof lies in the center of a
vast and largely unexplored application domain for automated and
interactive theorem provers.
